### 2024-09-04 Week 2 Journal - Grammar of Graphics/week2.qmd
Week 2 initial commit

---
format: 
  pdf:
    linestretch: 2
bibliography: references.bib
---

Tyler Wiederich | STAT 992 | [GitHub Repo](https://github.com/TWiedRW/stat-992)

# Prompt

Read the following introductory material on the grammar of graphics. In your 1-2 page reflection (see rubric for grading criteria), consider the scope of the grammar of graphics compared to your experience with plotting libraries in R (such as base graphics and ggplot2), SAS, and python.

Then, consider the low-level concepts of marks and channels discussed in Ch. 5 of Visualization Analysis and Design. How do these connect in to your understanding of the grammar of graphics?

# Response

The grammar of graphics outlines a process for which visualizations are created.